---
title: clustering and classification
author: Susanna Jernberg
output:
  html_document: default
---
#Clustering and classification

### Sucture of the data

In this exercise, I have used "Boston" dataset that is available in R package "Mass". It is a data about housing value in suburbs of Boston. Below I present the dimensions and sturcture of the dataset.

```{r}
library(MASS)
library(tidyverse)
library(corrplot)
data("Boston")
dim(Boston)
str(Boston)

```
### Summary and graphical overview of the data
```{r}
summary(Boston)
pairs(Boston)
```
Here I have looked more in detail about the correlations in the data
```{r}
cor_matrix<-cor(Boston)
cor_matrix%>%round(digits=2)
corrplot(cor_matrix, method="circle", type="upper", cl.bos="b", tl.pos="d", tl.cex=0.6)
```
### Data scaling
For scaling, the column means from the corresponding columns need to be subtraced and their difference is devided with standard deviation.
```{r}
boston_scaled<-scale(Boston)
summary(boston_scaled)
class(boston_scaled)
boston_scaled<-as.data.frame(boston_scaled)
class(boston_scaled)


```
Here I create a categorical variable of the crime rate in the Boston dataset 
```{r}
summary(Boston$crim)
bins<-quantile(boston_scaled$crim) #create a quantile vector of the crime in the boston_scaled
bins
classes<-c("low", "med_low", "med_high", "high")
classes
crime<-cut(boston_scaled$crim, breaks=bins, include.lowest = TRUE, labels=classes)
table(crime)
boston_scaled<-dplyr::select(boston_scaled, -crim)
boston_scaled<-data.frame(boston_scaled, crime)
```
Next I will divide the data to train and test set
```{r}
n<-nrow(boston_scaled)
ind<-sample(n, size= n*0.8) ##choose randomly 80 % of the rows
train<-boston_scaled[ind, ]
test<-boston_scaled[-ind,]
correct_classes<-test$crime
test<-dplyr::select(test, -crime) # remove the crime variable from test data
```
### Fitting the linear discriminant model
```{r}
lda.fit<-lda(crime~., data=train)
lda.fit
lda.arrows<-function(x, myscale=1, arrow_heads=0.1, color="red", tex=0.75, choices=c(1,2)){
  heads<-coef(x)
  arrows(x0=0, y=0,
         x1=myscale*heads[,choices[1]],
         y1=myscale*heads[,choices[2]], col=color, length=arrow_heads)
  text(myscale*heads[,choices], labels=row.names(heads), cex=tex, col=color, pos=3)
} ## function for lda biplot arrows
classes<-as.numeric(train$crime)
plot(lda.fit, dimen=2, col=classes, pch=classes)
lda.arrows(lda.fit, myscale=3)
```
###Prediction from the dataset
### Fitting the linear discriminant model
```{r}
lda.pred<-predict(lda.fit, newata=test) #this should continue with r code: table(correct=correct_classes, predicted=lda.pred$class) but for some reason my R doesn't allow to perform this and says only error "all arguments must have the same length"
```
### Distance measures
```{r}
dist_eu<-dist(Boston)
summary(dist_eu)
```
### K-means clustering
```{r}
km<-kmeans(Boston, centers=3)
pairs(Boston[6:10], col=km$cluster)
set.seed(123)
k_max<-10
twcss<-sapply(1:k_max, function(k){kmeans(Boston, k)$tot.withinss})
qplot(x=1:k_max, y=twcss, geom="line")
km<-kmeans(Boston, centers=2)
pairs(Boston, col=km$cluster)

```